Optimal Particle Paths Around Points in $\R^2$ with Constrained Acceleration

### 2013-09-29 cone_theorems.tex
Add cone proofs.

@@

Before we attempt to tackle the problem of finding the optimal path around a cone at constant velocity, we may want to find some intuition for the problem with a simple, but related problem. In particular, if a particle is traveling around a cone, we would like to know what circle it traces as it goes around the cone.

@@

To make a simple model of this problem, we will use a quadrilateral to model a path around a cone. We will construct a quadrilateral with particular constaints that models the constraints of a path around a cone, and we will try to find the parameters that minimize the perimeter around the quadrilateral. Let us call the sides of the quadrilateral $s_1, s_2, s_3$ and $s_4$. Imagine that $s_1$ and $s_2$ are on opposite sides of the quadrilateral with fixed lengths $l_1$ and $l_2$ respectively. Now we will constrain the problem to be related to the problem of a particle traveling around a cone: imagine $s_1$ and $s_2$ are connected by a bar $s_b$ of length $d$. The bar will be perpendicular to $s_1$. The angle that $s_b$ forms with $s_2$ will be called $\theta$. We will try to find the optimal $\theta$ that minimizes the perimeter around the quadrilateral.

\subsection{Symmetric Quadrilateral}

Let us begin by solving the simplest version of this problem. Let us imagine that $s_b$ is connected to the midpoints of $s_1$ and $s_2$. We know that the perimeter will be given by $l_1 + l_2 + l_3 + l_4$ where $l_3$ and $l_4$ are the lengths of the sides of $s_3$ and $s_4$, respectively. We are given $l_1$ and $l_2$, but we will need to compute $l_3$ and $l_4$ as functions of $l_1, l_2,d, $ and $\theta$.

We can find $l_3$ by using the fact that it forms a right triangle. We know that $l_3^2 = m^2 + n^2$. Finding $l_4$ is similar. Therefore, we just need to find $m$ and $n$. This can be done by using the fact that $m = \frac{l_1}{2} - \frac{l_2}{2} \sin \theta$. This is just the upper half of $s_1$ minus the projection of $s_2$ onto $s_1$. We can find $n$ similarly: $n = d - \frac{l_2}{2} \cos \theta$. This is just the bar $s_b$ minus the projection of $s_2$ onto the bar. Thus by substituting, we have:
\begin{eqnarray}
  l_3 = \sqrt{ \left(\frac{l_1}{2} - \frac{l_2}{2} \sin \theta \right)^2 + \left( d - \frac{l_2}{2} \cos \theta \right)^2 }
\end{eqnarray}

We can do a similar analysis on $l_4$, only remembering that $n$ for $l_4$ gets extended by the projection onto $s_4$ instead of shrunken. We therefore have:
\begin{eqnarray}
  l_4 = \sqrt{ \left(\frac{l_1}{2} - \frac{l_2}{2} \sin \theta \right)^2 + \left( d + \frac{l_2}{2} \cos \theta \right)^2 }
\end{eqnarray}

Now, to minimize the perimeter with respect to $\theta$, we want to minimize $l_1 + l_2 + l_3 + l_4$. Since we know that $l_1$ and $l_2$ are fixed, we really want to minimize $l_3 + l_4$ with respect to $\theta$. The other thing to note is that we're minimizing positive distances. We will invoke the following lemma so that we can simplify our expression for $\min l_3 + l_4$:

\begin{lemma}
  If $f(t), g(t) > 0$ and $k(t) > 0$ is a strictly monotonically increasing function for all $t$, then $\argmin_t k(f(t)) + k(g(t)) = \argmin_t f(t) + g(t)$.
  \label{lemma:minsqrt}
\end{lemma}
\proof Let $t_1, t_2$ be such that $f(t_1) + g(t_1) < f(t_2) + g(t_2)$. In this proof, we will show that $k(f(t_1)) + k(g(t_1)) < \sqrt{f(t_2)} + \sqrt{g(t_2)}$. Since $k(t)$ is a strictly monotonically increasing function when $t > 0$, we know that $k(x) < k(y)$ if and only if $x < y$ (assuming we can confine $x, y$ to be non-negative).

Because this is the case, we see that $k(f(x)) < k(f(y))$ if and only if $f(x) < f(y)$ (the same goes for $g$). Thus, we see that if we have found the minimum $t_m$ to $\argmin_t k(f(t)) + k(g(t))$, then it is the case that $k(f(t_m)) + k(g(t_m)) < k(f(t)) + k(g(t))$ for all $t \neq t_m$ (again where $t > 0$). Following our train of logic, we see that $f(t_m) + g(t_m) < f(t) + g(t)$ for all $t \neq t_m$, which means that $t_m$ is a minimum of $f(t) + g(t)$. Thus by finding a minimum $t_m$ to $k(f(t)) + k(g(t))$, we also found a minimum to $f(t) + g(t)$. \qed

Since we've proven this lemma, we can invoke it upon $\min l_3 + l_4$. Since $l_3 = \sqrt{z_3}$ and $l_4 = \sqrt{z_4}$, we can use $k(t) = \sqrt{t}$ and we can write $\min l_3 + l_4 = \min z_3 + z_4$ by using our lemma. Thus, we now want to solve the problem:
\begin{eqnarray}
  \argmin_{\theta} &2& \left(\frac{l_1}{2} - \frac{l_2}{2} \sin \theta \right)^2 \\
  &+& \left( d - \frac{l_2}{2} \cos \theta \right)^2 + \left( d + \frac{l_2}{2} \cos \theta \right)^2
\end{eqnarray}

Now, we can expand out our expression and use the fact that $\sin^2 \theta + \cos^2 \theta = 1$ to obtain a much simpler (but equivalent) minimization problem:
\begin{eqnarray}
  \argmin_{\theta} \frac{l_1^2}{2} + \frac{l_2^2}{2} + 2d^2 - l_1 l_2 \sin \theta
\end{eqnarray}

@@

Thus, we see that $\theta = \frac{\pi}{2}$ minimizes the perimeter of the quadrilateral in the symmetric case.

@@

Given a cone setup consisting of 3 cones at locations $X_A$, $X_B$, and $X_C$.

\includefig{cones_fix_s.eps}{A}{cones-fix-s}

\begin{align}
dB& = \sqrt{x_B + y_B}\\
\intertext{Applying the triangle inequality...}
d_c& = \alpha_c - \tan^{-1}\paren{\frac{y_B}{x_B}}\\
\end{align}

\includefig{cones_fix_s_2.eps}{A}{cones-fix-s-2}

\begin{align}
l_1& = \sqrt{d_c - r}\\
x_c = x_B - r \cos{\alpha_c}\\
y_c = y_B - r \sin{\alpha_c}\\
d_c& = \alpha_c - \tan^{-1}\paren{\frac{y_c}{x_c}}\\
\end{align}

@@

\proof

>>>>

Now we are finally able to tackle the model problem of optimal paths around cones. The cone problem we will solve will involve only 3 cones, but first we will solve a simpler but related problem of minimizing the path length around a quadrilateral.

@@

We would like to minimize the path length around the quadrilateral shown in Figure \ref{fig:quad}.

@@

We will constrain the problem such that it is similar to the problem of a particle traveling around a cone. Sides $\overline{BC}$, $\overline{DA}$, and $\overline{EF}$ have fixed lengths $l_1$, $l_2$, and $w$ respectively. Points E and F bisect $\overline{DA}$ and $\overline{BC}$ respectively. The goal is

\[
\argmin_{\theta} \bracket{l_{AB} + l_{BC} + l_{CD}}
\]

\begin{align}
  x_c& = w - \frac{l2}{2 \cos{\theta}}\\
  y_c& = \frac{l_2}{2} - \frac{l_2}{2 \sin{\theta}}\\
  l_{AB}& = \sqrt{x_c^2 + y_c^2}\\
  & = \sqrt{ \paren{\frac{l_1}{2} - \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }\\
  l_{BC}& = l2\\
  l_{CD}& = \sqrt{\paren{\frac{l_1}{2} + \frac{l_2}{2 \cos\theta}}^2 + \paren{w - \frac{l_2}{2 \sin\theta}}^2 }
\end{align}

Since we know that $l_{BC}$ is fixed the problem reduces to

\[
\argmin_{\theta} \bracket{l_{AB} + l_{BC} + l_{CD}}
\]

Because all distances are positive, we can use Lemma \ref{lemma:conv-argmin}, defined in the Appendix, to simplify our problem:

If we define $l_{AB} = \sqrt{z_{AB}}$ and $l_{CD} = \sqrt{z_{CD}}$, we can use \ref{lemma:conv-argmin} with $k(t) = \sqrt{t}$ to simplify our problem to

\begin{multline}
  \argmin_{\theta} \left[ 2\paren{\frac{l_1}{2} - \frac{l_2}{2 \sin\theta}}^2 \right.\\
  \left. + \paren{w - \frac{l_2}{2 \cos\theta}}^2 + \paren{w + \frac{l_2}{2 \cos\theta}}^2 \right]
\end{multline}

Now, we can expand out our expression and use Lemma the fact that $\sin^2 \theta + \cos^2 \theta = 1$ to obtain a much simpler (but equivalent) minimization problem:
\begin{eqnarray}
  \argmin_{\theta} \frac{l_1^2}{2} + \frac{l_2^2}{2} + 2d^2 - l_1 l_2 \sin \theta
\end{eqnarray}

@@

@@

Given a particle with zero tangential acceleration and bounded centripetal acceleration, $a_c \le a_{c,max}$, and a cone setup consisting of 3 cones, A, B, and C, at locations $\bvec{X_A}$, $\bvec{X_B}$, and $\bvec{X_C}$ as is shown in Figure \ref{fig:3-cones}.

\includefig{3_cones.eps}{3 Cone setup}{3-cones}

The optimal path [TODO]

@@

\begin{proof}
Since speed is constant, the optimal path will be the shortest path. The optimal path is of the form, tr. Thus, the problem reduces to a trigonometric problem.

\includefig{cones_fix_s.eps}{Position of center of turn}{cones-fix-s}

The turn must touch the cone B, otherwise the trajectory will either not go around the cone, or be longer than necessary. Therefor, the center of the turn must be a distance $r$ away from

\includefig{cones_fix_s_2.eps}{Path length}{cones-fix-s-2}

\begin{align*}
x_c = x_B - r \cos{\alpha_c}\\
y_c = y_B - r \sin{\alpha_c}\\
l_1& = \sqrt{d_c - r}\\
l_2& = r \paren{\alpha_c - \tan^{-1}\paren{\frac{y_c}{x_c}}}\\
\end{align*}
\end{proof}

### 2013-09-29 math_review.tex
Add more proofs.
Add cone proofs.

\subsection{Coordinate Systems\label{sec:notation}}

\subsubsection{Time Derivatives}

\begin{align*}
\dot{a}& = \frac{da}{dt}\\
%
\ddot{a}& = \frac{d^2a}{dt^2}
\end{align*}

@@

\begin{align}
R& = \abs{\frac{\paren{\dot{x}^2 + \dot{y}^2}^{3/2}}{\dot{x}\ddot{y} - \dot{y}{\ddot{x}}}}\\
%
&= \abs{\frac{s^2}{a_t}}\tag{for $a_c = $}
\end{align}

\subsection{Scalar Calculus in Polar Coordinates}

\begin{align}
s& = \frac{dl}{dt}\\
%
a_t& = \frac{ds}{dt}\\
\end{align}

@@

\begin{align}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
%
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
%
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align}

\subsection{Relations}

\begin{align}
l& = \int_{t=0}^{T} \norm{\boldsymbol{\vec{v}} \, } \; dt\\
%
s& = \norm{\boldsymbol{\vec{v}}}\\
%
a_t& = \norm{\boldsymbol{\vec{a}}} \cdot \boldsymbol{\hat{v}} = \norm{\boldsymbol{\vec{a}}} \cdot \frac{\boldsymbol{\vec{v}}}{\norm{\boldsymbol{\vec{v}}}}\\
%
a_c& = \norm{\boldsymbol{\vec{a}}} \times \boldsymbol{\hat{v}} = \norm{\boldsymbol{\vec{a}}} \times \frac{\boldsymbol{\vec{v}}}{\norm{\boldsymbol{\vec{v}}}}
\end{align}

>>>>

\subsection{Analysis}

\begin{lemma}\label{lemma:conv-argmin}
  If $f(t), g(t) > 0$ and $k(t) > 0$ is a strictly monotonically increasing function for all $t$, then $\argmin_t k(f(t)) + k(g(t)) = \argmin_t f(t) + g(t)$.
  \label{lemma:minsqrt}
\end{lemma}
\begin{proof}
Let $t_1, t_2$ be such that $f(t_1) + g(t_1) < f(t_2) + g(t_2)$. In this proof, we will show that $k(f(t_1)) + k(g(t_1)) < \sqrt{f(t_2)} + \sqrt{g(t_2)}$. Since $k(t)$ is a strictly monotonically increasing function when $t > 0$, we know that $k(x) < k(y)$ if and only if $x < y$ (assuming we can confine $x, y$ to be non-negative).

Because this is the case, we see that $k(f(x)) < k(f(y))$ if and only if $f(x) < f(y)$ (the same goes for $g$). Thus, we see that if we have found the minimum $t_m$ to $\argmin_t k(f(t)) + k(g(t))$, then it is the case that $k(f(t_m)) + k(g(t_m)) < k(f(t)) + k(g(t))$ for all $t \neq t_m$ (again where $t > 0$). Following our train of logic, we see that $f(t_m) + g(t_m) < f(t) + g(t)$ for all $t \neq t_m$, which means that $t_m$ is a minimum of $f(t) + g(t)$. Thus by finding a minimum $t_m$ to $k(f(t)) + k(g(t))$, we also found a minimum to $f(t) + g(t)$.
\end{proof}

\subsection{Coordinate Systems}

@@

For $a_t = 0$

\[
R= \abs{\frac{v^2}{a_c}}
\]

@@

\begin{align*}
\boldsymbol{x}& = r \boldsymbol{\hat{r}}\\
\boldsymbol{\vec{v}}& = \dot{r} \boldsymbol{\hat{r}} + r \dot{\phi} \boldsymbol{\hat{\phi}}\\
\label{eq:a_vec_def}\boldsymbol{\vec{a}}& = \paren{\ddot{r} - r \dot{\phi}^2} \boldsymbol{\hat{r}} + \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}} \boldsymbol{\hat{\phi}}
\end{align*}

### 2013-09-29 paper.tex
Add more proofs.
Add cone proofs.

\title[The 18.821 report]{The 18.821 Mathematics Project Lab Report 
[Proofs]} 
% the first [...] gives a brief version of the title, which is long!

@@

@@

@@

\bibliographystyle{plain}
\begin{thebibliography}{9}
\bibitem{wiki_pcoords}
\url{http://en.wikipedia.org/wiki/Polar_coordinate_system}
\end{thebibliography}

@@

From Spring 2011 or before; revised by Miller, Spring 2013.

>>>>

\title[Optimal Paths in R2]{Optimal Particle Paths Around Points in $\R^2$ with Constrained Acceleration}

@@

\subsection*{Abstract}

Optimal particle paths in $\R^2$ between points and around points are investigated in this paper. A paramerization of a particle's motion around a point is defined. From this parameterization, we derive a general solution for an optimal path with constant speed between 2 points. The results of this paper can by applied towards calculating optimal object trajectories in physics when acceleration is constrained.

@@

\section{Conclusion}
\input{conclusion.tex}

@@

\input{bibliography.tex}

@@

### 2013-09-29 particle_definitions.tex
Add cone proofs.

In this section, we will provide some basic definitions about particles and paths. These will lay the groundwork for thinking about optimal paths. We will begin by defining the description of a particle and then define various types of paths along which a particle can travel.

@@

\begin{definition}
  A n-dimensional particle, $p$, is an object with zero volume that travels along a n-dimensional path. The particle may have conditions on its position, velocity, and acceleration in $\R^n$.
\end{definition}

\begin{definition}
  A valid path $\gamma(t)$ for a particle $p$ is a path such that all conditions on the particle are satisfied at every point along the path.
\end{definition}

\begin{definition}
  A path between two points, $\bvec{X_1}$ and $\bvec{X_2}$ is a path, $\gamma(t)$ where $\gamma(0) = \bvec{X_1}$ and $\gamma(T_{f, \gamma}) = \bvec{X_2}$.
\end{definition}

@@

  A particle's speed is $v$, and its direction of motion is $\bhat{v}$

@@

\begin{definition}
  The centripetal acceleration, $\bvec{a_c}$, of a particle, $p$, is the component of the acceleration of $p$ perpendicular to its direction of motion.

  In rectangular coordinates, the sign of $a_c$ is the projection of $\bhat{a_c}$ onto $\bhat{x}$.

  In polar coordinates, the sign of $a_c$ is difined as the sign of the projection of $\bhat{a_c}$ onto $\bhat{r}$.
\end{definition}

\begin{definition}
  The tangential acceleration, $\bvec{a_t}$, of a particle, $p$, is the component of the acceleration of $p$ in its direction of motion, $\bhat{v}$. $a_t = \frac{ds}{dt}$.
\end{definition}

@@

The motion of the 2-dimensional particles in this paper will be described in the polar coordinate system, shown in Figure \ref{fig:polar-param}.

@@

\begin{lemma}
The time derivative of $\theta$ is given by
\[
\frac{d\beta}{dt} = \frac{a_c}{s}
\]
\end{lemma}

@@

So...
\[
\frac{d\beta}{dt} = \frac{a_c}{v}
\]
\end{proof}

The proof of the lemma in the case where $a_t$ is also nonzero is very similar, since the component of $\bvec{dv}$ in the direction $\bhat{v}$ is negligeable compared to $v$.

@@

\begin{align}
  \frac{dr(t)}{dt}& = -s(t) \, \cos\paren{\theta(t)}\label{eq:r-deriv}\\
  \frac{d\phi(t)}{dt}& = s(t) \, \sin\paren{\theta(t)}\label{eq:phi-deriv}\\
  \frac{d\theta(t)}{dt}& = \frac{d\phi(t)}{dt} - \frac{a_c(t)}{s(t)}\\
  &= s(t) \, \sin\paren{\theta(t)} - \frac{a_c(t)}{s(t)}\label{eq:theta-deriv}\\
  \intertext{Applying the chain rule to (\ref{eq:r-deriv})}
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} + s(t) \sin\paren{\theta(t)} \frac{d\theta(t)}{dt}\\
  &\begin{aligned}
    = -\frac{ds(t)}{dt} \cos\paren{\theta(t)} &+ s(t)^2 \sin^2\paren{\theta(t)}\\
    &\qquad - s(t) \sin\paren{\theta(t)} \frac{a_c(t)}{s(t)}\label{eq:r-deriv-2}
  \end{aligned}
\end{align}

@@

  1. The functions $\phi(t)$ and $\theta(t)$ are continuous.

  2. For two times $t_1$ and $t_2$, s.t. $t_1 < t_2$, $\theta(t_1) > a$ and $\theta(t_2) < a$, then $\theta(t_c) = a$ for some $t_c \in [t_1, t_2]$.
\end{lemma}

\begin{proof}
First off, it should be noted that a solution to \ref{eq:theta-deriv} exists because the right hand side is Lipschitz continuous. The proof of 1.\ follows directly from the fact that the derivative of $\theta(t)$ exist and is bounded. 2.\ is just a restatement of the intermediate value theorem.
\end{proof}

\begin{lemma}
  For a particle, $p$, with nonzero speed, and no centripetal acceleration for $t \ge t_0$, then
  \[
    \begin{cases}
      \theta(t) \to \pi \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(t_0) > 0\\
      \theta(t) = 0 \quad \text{for} \quad t \ge t_0 \qquad &\text{if} \;\; \theta(t_0) = 0
    \end{cases}
  \]
\end{lemma}

@@

For the first case

@@

\begin{align*}
\frac{d\theta(t)}{dt} = 0 \qquad \text{for} \;\; t \ge t_0\\
\intertext{So...}
\theta(t) = 0 \qquad \text{for} \;\; t \ge t_0
\end{align*}

>>>>

In this section, we will provide some basic definitions about particles and paths. These will lay the groundwork for thinking about optimal paths.

@@

\begin{definition}
  A n-dimensional particle, $p$, is an object with zero volume that travels along a n-dimensional path. Traveling along a path means that the particle is at position $\lambda(t)$ at time $t$, for all $t \in [0, T_{f,\lambda}]$. The particle may have conditions on its position, velocity, and acceleration in $\R^n$.
\end{definition}

\begin{definition}
  A valid path $\gamma(t)$ for a particle $p$ is a path along which $p$ can travel, such that all its conditions are satisfied at every point along the path.
\end{definition}

\begin{definition}
  A path between two points, $\bvec{X_1}$ and $\bvec{X_2}$ is a path, $\gamma(t)$, where $\gamma(0) = \bvec{X_1}$ and $\gamma(T_{f, \gamma}) = \bvec{X_2}$.
\end{definition}

@@

  Furthermore, a particle's speed is $v$, and its direction of motion is $\bhat{v}$

@@

\begin{definition}
  The centripetal acceleration, $\bvec{a_c}$, of a particle, $p$, is the component of its acceleration in the direction perpendicular to its direction of motion.

  In rectangular coordinates, the sign of $a_c$ is defined to be the sign of the projection of $\bhat{a_c}$ onto $\bhat{x}$.

  In polar coordinates, the sign of $a_c$ is defined to be the sign of the projection of $\bhat{a_c}$ onto $\bhat{r}$.
\end{definition}

\begin{definition}
  The tangential acceleration, $\bvec{a_t}$, of a particle, $p$, is the component of the acceleration of $p$ in its direction of motion, $\bhat{v}$.

  \[
  a_t = \frac{dv}{dt}
  \]
\end{definition}

@@

Unless otherwise specified, the motion of 2-dimensional particles in this paper will be described in polar coordinates, along with an extra parameter $\theta$ as is shown in Figure \ref{fig:polar-param}.

@@

$\theta(t), \phi(t) \in [-\pi, \pi]$. From now on, $\abs{\theta}$ and $\abs{\phi}$ will be used, since there is a symmetry in the system about $\bhat{r}$.

\begin{lemma}
The time derivative of $\theta$ is given by
\[
\frac{d\abs{\theta(t)}}{dt} = \frac{a_c}{v}
\]
\end{lemma}

@@

Since $dv = d\abs{\theta} v$, then
\[
\frac{d\abs{\theta}}{dt} = \frac{a_c}{v}
\]
\end{proof}

The proof of the lemma in the case where $a_t$ is also nonzero is very similar, since the component of $\bvec{dv}$ in the direction $\bhat{v}$ is negligible compared to $v$.

@@

\begin{align}
  \frac{dr(t)}{dt}& = -v(t) \, \cos\paren{\abs{\theta(t)}}\label{eq:r-deriv}\\
  \frac{d\abs{\phi(t)}}{dt}& = \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}}\label{eq:phi-deriv}\\
  \frac{d\abs{\theta(t)}}{dt}& = \frac{d\abs{\phi(t)}}{dt} - \frac{a_c(t)}{v(t)}\\
  &= \frac{v(t)}{r(t)} \, \sin\paren{\abs{\theta(t)}} - \frac{a_c(t)}{v(t)}\label{eq:theta-deriv}\\
  \intertext{Applying the chain rule to (\ref{eq:r-deriv})}
  \frac{d}{dt} \frac{dr(t)}{dt}& = -\frac{dv(t)}{dt} \cos\paren{\abs{\theta(t)}} + v(t) \sin\paren{\abs{\theta(t)}} \frac{d\abs{\theta(t)}}{dt}\\
  &\begin{aligned}
    = -\frac{dv(t)}{dt} \cos\paren{\abs{\theta(t)}} &+ \frac{v(t)^2}{r(t)} \sin^2\paren{\abs{\theta(t)}}\\
    &\qquad - v(t) \sin\paren{\abs{\theta(t)}} \frac{a_c(t)}{v(t)}\label{eq:r-deriv-2}
  \end{aligned}
\end{align}

@@

  1. The function $\abs{\theta(t)}$ is continuous.

  2. Given $t_1, t_2 \in \R^2$, s.t. $t_1 < t_2$, $\theta(t_1) > a$ and $\theta(t_2) < a$, then $\theta(t_c) = a$ for some $t_c \in [t_1, t_2]$.
\end{lemma}

\begin{proof}
First off, it should be noted that a solution to (\ref{eq:theta-deriv}) exists because the right hand side is Lipschitz continuous. The proof of 1.\ follows directly from the fact that the derivative of $\abs{\theta(t)}$ exists and is bounded. 2.\ is just a restatement of the intermediate value theorem.
\end{proof}

\begin{lemma}\label{lemma:polar-assymptote}
  For a particle, $p$, with nonzero speed, and zero centripetal acceleration for $t \ge t_0$, then
  \[
    \begin{cases}
      \abs{\theta(t)} \to \pi \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(t_0) > 0\\
      \abs{\theta(t)} = 0 \quad \text{for all} \quad t \ge t_0 \qquad &\text{if} \;\; \theta(t_0) = 0
    \end{cases}
  \]

    Furthermore,

  \[
    \begin{cases}
      r(t) \to \infty \quad \text{as} \quad t \to \infty \qquad &\text{if} \;\; \theta(t_0) > 0\\
      r(t) \to 0 \quad \text{as} \quad t \ge t_0 \qquad &\text{if} \;\; \theta(t_0) = 0
    \end{cases}
  \]
\end{lemma}

@@

For the first case, $\frac{d\abs{\theta(t)}}{dt} > 0$, except when $\abs{\theta(t)} = 0, \pi$, which means that $\theta(t)$ is monotonically increasing. Since $\phi(t)$ is bounded above py $\pi$, $\phi(t) \to \pi \;\; \text{as} \;\; \infty$.

@@

\begin{align*}
\frac{d\abs{\theta(t)}}{dt} = 0 \qquad \text{for} \;\; t \ge t_0\\
\intertext{So}
\abs{\theta(t)} = 0 \qquad \text{for} \;\; t \ge t_0
\end{align*}

[TODO]

### 2013-09-29 path_theorems.tex
Add cone proofs.

\begin{lemma}
If $a_t=0$ and $\norm{a_c} \le a_{c,max}$, then the minimum radius of curvature of a point trajectory is given by
\begin{equation}
R_{min} = \frac{v^2}{a_{c,max}}
\end{equation}
\end{lemma}

\begin{theorem}
For a particle, $p$, traveling along an optimal path, $\hat{\gamma}(t)$ between two points $X_1$ and $X_2$, 
\end{theorem}

\begin{lemma}
Every optimal trajectory is constructed from line segments and circular sections of radius $R_{min}$. Every point along an optimal trajectory where $a_t=0$ has $R = \infty$ or $R = R_{min}$.
\end{lemma}

\proof

If we look at a position, p, on a trajectory. We can align a rectangular coordinate system with this point, such that $\hat{y} = \hat{v}$.

\begin{align}
a_c& = \ddot{r} - r \dot{\phi}^2\\
%
a_t& = \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}}\\
%
0& = \frac{1}{r} \frac{d}{dt} \paren{r^2 \dot{\phi}}\\
%
r^2 \dot{\phi}& = constant
\end{align}

\qed

@@

\begin{theorem}
  Given points $(x_1, y_1), (x_2, y_2) \in \mathrm{R}^2$ and a particle $p$ whose initial position is $(x_1, y_1)$ which moves with acceleration bounded by $\bar{a}$, the fastest path $\hat{\gamma}(t)$ which $p$ can trace from $(x_1, y_1)$ to $(x_2, y_2)$ follows the straight line where all coordinates $(x,y)$ on the straight line are given by:
  \begin{eqnarray}
    y = \frac{y_2 - y_1}{x_2 - x_1} x + y_1
  \end{eqnarray}
\end{theorem}
\proof Let's transform the problem. We can reset our coordinate axes so that $(x_1, y_1)$ is set to the origin and $(x_2, y_2)$ is on the x-axis. In this new coordinate system, we have transformed the following:
\begin{eqnarray}
  (x_1, y_1) &\to& (0,0) \\
  (x_2, y_2) &\to& (x_2', 0)
\end{eqnarray}

For convenience of notation, we will now refer to $x_2'$ as $x_2$.

Now let us examine the particle's motion in the $x$ direction. Let $a_t(t)$ be the tangential acceleration at time $t$ in the $x$ direction. Then we can obtain the speed of the particle $s(t)$ at time $t$ in the $x$ direction like so:
\begin{eqnarray}
  s(t) = \int_0^t a_t(t_1) dt_1
\end{eqnarray}

To find the distance $d(t)$ travelled up to time $t$ in the $x$ direction, we can use the relation:
\begin{eqnarray}
  d(t) &=& \int_0^t s(t_2) dt_2 \\
       &=& \int_0^t \int_0^t a_t(t_1) dt_1 dt_2
\end{eqnarray}

Recall that the acceleration of the point mass $p$ is bounded by $\bar{a}$. This means that $a_t(t) \leq \bar{a}$ for all $t$. Therefore, we see:
\begin{eqnarray}
  d(t) &\leq& \int_0^t \int_0^t \bar{a} dt_1 dt_2 \\
       &=& \frac{\bar{a} t^2}{2}
\end{eqnarray}

@@

Now, we have seen that the fastest path in the transformed coordinates travels exactly on the $x$ axis so that $y = 0$ anywhere along the fastest path. Notice, however, that the $x$ axis in the transformed coordinates is given exactly by the following line:
\begin{eqnarray}
  y = \frac{y_2 - y_1}{x_2 - x_1} x + y_1
\end{eqnarray}

Thus, we see that the fastest path in the original coordinates follows the above equation, which is what we wanted to show.
\qed

>>>>

The simplest problem to tackle first, is the optimal trajectory from a current position, $\bvec{X_0}$ to a final position, $\bvec{X_f}$. Working in a polar coordinate system centered on $X_f$, it is clear, that our goal is achieved when $r = 0$. Furthermore, $\theta(t)$ must be $0$ before reaching $X_f$, otherwise, it will grow exponentially fast as $r(t) \to 0$, which can be seen in equation (\ref{eq:r-deriv}).

\begin{lemma}
If $a_t=0$ and $a_c \le a_{c,max}$, then the minimum radius of curvature of a particle's trajectory is given by
\begin{equation}
R_{min} = \frac{v^2}{a_{c,max}}
\end{equation}
\end{lemma}

\begin{proof}
This comes directly from the definition of radius of curvature

\[
R = \frac{v^2}{a_c}
\]

and the fact that

\[
a_c \le a_{c,max} \to \frac{1}{a_c} \ge \frac{1}{a_{c,max}}
\]
\end{proof}

\begin{theorem}
Given a 2-dimensional particle, $p$, with bounded centripetal acceleration, $a_c \le a_{c,max}$ and zero tangential acceleration, every valid optimal path, $\hat{\gamma}(t)$ between two points $X_1$ and $X_2$, has a radius of curvature of $0$ or $a_{c,max}$ at every point along it.
\end{theorem}

\begin{proof}
[TODO]
\end{proof}

\begin{corollary}
Given a 2-dimensional particle, $p$, with bounded centripetal acceleration, $a_c \le a_{c,max}$ and zero tangential acceleration, every valid optimal trajectory is constructed from line segments and circular sections of radius $R_{min}$, that are tangent at their intersections.
\end{corollary}

\begin{proof}
[TODO]
\end{proof}

@@

\begin{theorem}
  Given a 2-dimensional particle $p$ whose initial position is $\bvec{X_1}$ and moves with bounded tangential acceleration, $a_t \le a_{t,max}$, and infinite centripetal acceleration, the fastest path $\hat{\gamma}(t)$ which $p$ can trace from $\bvec{X_1}$ to $\bvec{X_2}$ is the line segment from $\bvec{X_1}$ to $\bvec{X_2}$:
\end{theorem}

\begin{proof}

Without loss of generality, we can define a cartesian coordinate system, where the origin is at $\bvec{X_1}$ and the positive x-axis passes through $\bvec{X_2}$.

Now let us examine the particle's motion in the $\bhat{x}$ direction. The speed of the particle is the following
\begin{equation}
  v_x(t) = \int_0^t a_x(t_1) dt_1
\end{equation}

To find the distance $l_x(t)$ travelled in the $\bhat{x}$ direction, we can use the relation:
\begin{align}
  d(t) &= \int_0^t v_x(t_2) dt_2 \\
  &= \int_0^t \int_0^t a_x(t_1) dt_1 dt_2
\end{align}

Recall that the tangential acceleration of the point mass $p$ is bounded by $a_{t,max}$. This means that $a_t(t) \leq \bar{a}$ for all $t$. Therefore, we see:
\begin{align}
  d(t) &\leq \int_0^t \int_0^t \bar{a} dt_1 dt_2\\
  &= \frac{\bar{a} t^2}{2}
\end{align}

@@

Now, we have seen that the fastest path in the transformed coordinates travels exactly on the $x$ axis so that $y = 0$ anywhere along the fastest path. Notice, however, that the $x$ axis in the transformed coordinates is given exactly by the following line:
\begin{equation}
  y = \frac{y_2 - y_1}{x_2 - x_1} x + y_1
\end{equation}

Thus, we see that the fastest path in the original coordinates follows the above equation, which is what we wanted to show.
\end{proof}
